---
title: "Class 8 Exercises"
output: html_document
date: "2024-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
```

## Exercises

### Easy

Do all the easy exercises from Chapter 9: **9E1** through **9E6**.


**9E1.** Which of the following is a requirement of the *simple* Metropolis algorithm?
(1) The parameters must be discrete
(2) The likelihood function must be Gaussian.
*(3) The proposal distribution must be symmetric. -> This one*

The simple Metropolis algorithm requires the proposals to arise from a symmetric distribution - spørg lige chris 

"The Metropolis algorithm works whenever the probability of proposing a jump to B from A is equal to the probability of proposing A from B, when the  proposal distribution is symmetric." p. 273 

**9E2.** Gibbs sampling is more efficient than the Metropolis algorithm. How does it achieve this extra efficiency? Are there any limitations to the Gibbs sampling strategy?

It uses *adaptive proposals*, meaning that it uses different *conjugate pairs*. Conjugate pairs are, I imagine, kinda like puzzle pieces. If a prior and a likelihood fits together in a particular way, then the posterior has the same structure as the prior (Chris explanation) p. 273 

Limitations to Gibbs sampling: 
- They can get stuck for long time in small regions of the posterior 
- Especially an issue with more parameters, as it's more likely for some parameters to be highly correlated (high correlation = narrow ridges, I imagine a mountain peak)
- "both Metropolis and Gibbs make too many dumb proposals of where to
go next. So they get stuck." p. 724 

**9E3.** Which sort of parameters can Hamiltonian Monte Carlo not handle? Can you explain why?

Can't handle discrete values. It needs to be able to 'stop' at any point 


### Medium

#### 9M1

```{r}

```

#### 9M2


```{r}

```



### Hard

#### 9H1

```{r}

# Your solution here

```

#### 9H2

```{r}

# Your solution herre

```
